Parametricity for Primitive Nested Types and GADTs

### 2020-12-16 free-theorems-nesteds/20-mscs.tex
finish body of paper draft except GADTs

Definition~\ref{def:wftypes} allows the formation of all of the types
from Section~\ref{sec:intro}:
\[\begin{array}{lll}
\mathit{List}\, \alpha & = & \mu \beta. \,\onet + \alpha \times
\beta\; = \; (\mu \phi. \lambda \beta.\,\onet + \beta \times \phi
\beta)\,\alpha\\ 
\mathit{PTree}\,\alpha & = & (\mu \phi. \lambda \beta.\,\beta +
\phi\,(\beta \times \beta))\,\alpha\\
\mathit{Forest}\,\alpha & = & (\mu \phi. \lambda \beta. \,\onet +
\beta \times \mathit{PTree}\,(\phi \beta))\,\alpha\\ 
\mathit{Bush}\,\alpha & = & (\mu \phi.\lambda \beta. \,\onet + \beta
\times \phi\,(\phi\beta))\,\alpha 
\end{array}\]
The representation of $\mathit{List}\,\alpha$ as the ADT $\mu
\beta. \onet + \alpha \times \beta$ cannot be functorial in
$\alpha$. By contrast, its representation as the nested type $(\mu
\phi. \lambda \beta.\,\onet + \beta \times \phi \beta)\,\alpha$ can be
considered either functorial in $\alpha$ or not, according to whether
$\alpha$ is in the functorial or non-functorial context in which this
type is formed. If $\emptyset;\alpha \vdash \mathit{List} \,\alpha$,
then $\vdash \Nat^\alpha \onet\, (\mathit{List}\,\alpha)$ is
well-formed; if $\alpha;\emptyset \vdash \List \,\alpha$, then it is
not.  Similarly, if $\mathit{Tree}\,\alpha\,\gamma = \mu \,\beta.\,
\alpha + \beta \times \gamma \times \beta$, then $\gamma;\emptyset
\vdash \Nat^\alpha (\mathit{List}\, \alpha) \, (\mathit{Tree} \,
\alpha\, \gamma)$ is well-formed, but $\emptyset;\gamma \vdash
\Nat^\alpha (\mathit{List}\, \alpha) \, (\mathit{Tree} \, \alpha\,
\gamma)$ is not. The body $F$ of a type $(\mu \phi.\lambda
\ol\alpha.F)\ol G$ can only be functorial in $\phi$ and the variables
in $\ol\alpha$. As a result, both $\phi$ and the variables in
$\ol\alpha$ must be non-functorial in $\mathit{GRose}\, \phi \, \alpha
= \mu\beta . \alpha \times \phi \beta$, $\mathit{GRose}\,\phi\,\alpha$
cannot be the (co)domain of a $\Nat$-type binding any variables, and
no type of the form $(\mu \phi. \lambda
\alpha. \mathit{GRose}\,\phi\,\alpha)\tau$ is well-formed in our
system. These restrictions ensure that the set and relational
interpretations of types in Section~\ref{sec:type-interp} are
functors. In Section~\ref{sec:conclusion} we discuss how more
expressive types might be accommodated.

@@

\lambda d\,\ol{B}\,
{\color{red}\ol{C}}.\,(\mathit{in}_{{T}^X_{H,\rho{\color{red}[\ol{\gamma :=
        C}]}}})_{\ol{B}} \;\;\mbox{ where } X \mbox{ is } \set \mbox{ when } \\ 

@@

\lambda d\,\eta\,\ol{B}{\color{red}\,\ol{C}}.\,
(\mathit{fold}_{T^X_{H,\rho{\color{red}[\ol{\gamma := C}]}}} \, (\lambda
\ol{A}.\,\eta_{\ol{A}\,\ol{C}}))_{\ol{B}}\\ 

@@

\section{Free Theorems for Nested Types}\label{sec:free-theorems}

@@

If $\Gamma; \ol{\alpha} \vdash F$ then the {\em identity} $\id_F$ on
$F$ is defined by $\id_F = \Gamma;\emptyset~|~\emptyset \vdash
L_{\ol{\alpha}}x.x : \Nat^{\ol{\alpha}} F\,F$, and if $\Gamma;
\emptyset \,|\, \Delta \vdash t: \Nat^{\overline{\alpha}} F\,G$ and
$\Gamma; \emptyset \,|\, \Delta \vdash s: \Nat^{\overline{\alpha}}
G\,H$ are terms then the {\em composition} $s \circ t$ of $t$ and $s$
is defined by $s \circ t = \Gamma; \emptyset\,|\, \Delta \vdash
L_{\overline{\alpha}} x. s_{\overline{\alpha}}(t_{\overline{\alpha}}x):
\Nat^{\overline{\alpha}} F\,H$. Then \[\setsem{\Gamma; \emptyset \,|\,
  \emptyset \vdash \id_{F} : \Nat^{\ol{\alpha}} F\,F} \rho\, \ast =
\id_{\lambda \ol{A}. \setsem{\Gamma; \ol{\alpha} \vdash F} \rho
  [\ol{\alpha := A}]}\] for any set environment $\rho$, and
\[\setsem{\Gamma; \emptyset \,|\, \Delta \vdash s \circ t:
  \Nat^{\overline{\alpha}} F\,H} = \setsem{\Gamma; \emptyset \,|\,
  \Delta \vdash s: \Nat^{\overline{\alpha}} G\,H} \circ
\setsem{\Gamma; \emptyset \,|\, \Delta \vdash t:
  \Nat^{\overline{\alpha}} F\,G}\]
Straightforward calculation shows that terms of $\Nat$-type behave as
natural transformations with respect to their source and target
functorial types, i.e., that
\begin{thm}\label{eq:ft-from-nat} 
  If\, $\Gamma; \emptyset \,|\, \Delta \vdash s : \Nat^{\overline{\alpha},
  \overline{\gamma}} F\,G$ and $\overline{\Gamma; \emptyset \,|\,
  \Delta \vdash t : \Nat^{\overline{\gamma}} K\, H}$, then
  \[\begin{array}{l}
  \hspace*{0.135in}
\setsem{\Gamma; \emptyset\,|\, \Delta
  \vdash 
  ((\map_G^{\overline{K}, \overline{H}})_\emptyset \,\overline{t}) \circ
(L_{\overline{\gamma}} z. s_{\overline{K}, \overline{\gamma}}
  z)
  : \Nat^{\overline{\gamma}} F[\overline{\alpha := K}]\,
  G[\overline{\alpha := H}]}\\
= \setsem{ \Gamma; \emptyset \,|\, \Delta \vdash
(L_{\overline{\gamma}} z.
  s_{\overline{H}, \overline{\gamma}} z)\circ
  ((\map_F^{\overline{K}, \overline{H}})_\emptyset \,
  \overline{t})  : \Nat^{\overline{\gamma}} F[\overline{\alpha :=
      K}]\, G[\overline{\alpha := H}]}
\end{array}\]
\end{thm}
\noindent

@@

\item
$\underline{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
  \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
    {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
  \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}$\; To
  see that if $d : \setsem{\Gamma;\emptyset \vdash \emptyset} \rho$
  then $$\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,
  \rho\,d$$ 
  is in $\setsem{\Gamma;\emptyset \vdash
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,
  \rho$, we first note that, for all $\ol{B}$ {\color{red} and $\ol{C}$},
  \begin{align*}
    &(\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,
  \rho\,d)_{\ol{B}{\color{red}\,\ol{C}}}\, \\
    =\,
    &(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}})_{\ol{B}}
  \end{align*}
  maps
  \begin{align*}
    &\setsem{\Gamma;\ol{\beta}{\color{red},\ol{\gamma}} \vdash H[\phi := (\mu
      \phi.\lambda {\overline \alpha}.H){\overline \beta}][\ol{\alpha
        := \beta}]}\rho[\ol{\beta := B}]{\color{red}[\ol{\gamma := C}]} 
        \\ = \,
    &T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}\, (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma
        := C}]}}) \, \ol{B}
  \end{align*}
        to $\setsem{\Gamma;\ol{\beta},{\color{red}\ol{\gamma}}
    \vdash (\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}
  \rho[\ol{\beta := B}]{\color{red}[\ol{\gamma := C}]} = (\mu
  T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}) \, \ol{B}$.  \\
  Secondly, we observe
  that
  \begin{align*}
    &\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline
      \beta}}\,\rho\,d \\ 
      = \, &\lambda
  \ol{B}{\color{red}\,\ol{C}}.\,(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma :=
          C}]}}})_{\ol{B}}
  \end{align*}
  is natural in $\ol{B}$ {\color{red} and $\ol{C}$}, since
  naturality of $\mathit{in}$ with respect to its functor argument and
  naturality of $\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}}$ ensure
  that the following diagram commutes for all $\ol{f : B \to B'}$
  {\color{red} and
  $\ol{g : C \to C'}$}:

  {\tiny
  \[\hspace*{-0.15in}\begin{tikzcd}[column sep=2.5in, row sep=0.75in]
%T^\set_{H,\rho[\ol{\gamma := C}]}\, (\mu T^\set_{H,\rho[\ol{\gamma := C}]})\, \ol{B}
%\ar[d, "{T^\set_{H,\id_\rho[\ol{\gamma := g}]}\,(\mu
%    T^\set_{H,\id_\rho[\ol{\gamma := g}]}) \, \ol{B}}"] \ar[r,
%  "{(\mathit{in}_{T^\set_{H,\rho[\ol{\gamma := C}]}})_{\ol{B}}}" ]
%& (\mu T^\set_{H,\rho[\ol{\gamma := C}]})\, \ol{B} \ar[d, "{(\mu
%    T^\set_{H,\id_\rho[\ol{\gamma := g}]}) \, \ol{B}}"]\\
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}\, (\mu
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\, \ol{B} \ar[d, "{T^\set_{H,\rho{\color{red}[\ol{\gamma :=
          C'}]}}\, (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\, \ol{f}}"]
\ar[r,"{(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})_{\ol{B}}}}" ] & 
  (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\, \ol{B}
\ar[d,"{\mu {T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}}\, {\ol{f}}}" ] 
\\
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}\, (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\,
\ol{B'} \ar[r, "{(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma :=
            C'}]}}})_{\ol{B'}}}"] & (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\,
\ol{B'}
\end{tikzcd}\]
}
  That, for all $\rho : \setenv$ and $d :
\setsem{\Gamma;\emptyset \vdash \emptyset}\rho$,
\[\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash
  \tin_H : \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
    {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
  \;(\mu \phi.\lambda {\overline \alpha}.H){\overline
    \beta}}\,\rho\,d\] satisfies the additional property needed for
it to be in 
\[\setsem{\Gamma;\emptyset \vdash
  \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
    {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
  \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,\rho\]
follows from the fact that, for every $\ol{R : \rel(B,B')}$
{\color{red} and $\ol{S : \rel(C,C')}$},
\[\hspace*{-0.2in}
  \begin{array}{ll}
 & (\,(\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
      \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
        {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
      \;(\mu \phi.\lambda {\overline \alpha}.H){\overline
        \beta}}\,\rho\,d)_{\ol{B}{\color{red},\ol{C}}},\,\\ & \hspace*{0.5in}(\setsem{\Gamma;\emptyset
      \,|\, \emptyset \vdash \tin_H : \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \,
      H[\phi := (\mu \phi.\lambda {\overline \alpha}.H){\overline
          \beta}][\ol{\alpha := \beta}] \;(\mu \phi.\lambda {\overline
        \alpha}.H){\overline
        \beta}}\,\rho\,d)_{\ol{B'}{\color{red},\ol{C'}}}\,)\\ &= (\,
    (\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}})_{\ol{B}},
    (\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}})_{\ol{B'}}\,)
\end{array}\]
has type
\[\begin{array}{ll}
& (\, T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}\, (\mu
T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}) \, 
\ol{B} \to (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}) \, \ol{B}, \, \\
& \hspace*{0.5in}T^\set_{H,\rho{\color{red} [\ol{\gamma := C'}]}}\, (\mu
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}) \, 
\ol{B'} \to (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}) \,\ol{B'} \, )\\
= &
 \relsem{\Gamma;\ol{\beta},{\color{red}\ol{\gamma}} \vdash H[\phi := (\mu
    \phi.\lambda {\overline \alpha}.H){\overline \beta}][\ol{\alpha :=
      \beta}]}\Eq_\rho[\ol{\beta := R}]{\color{red}[\ol{\gamma := S}]} \to\\
 & \hspace*{0.5in} \relsem{\Gamma;\ol{\beta}{\color{red},\ol{\gamma}} \vdash (\mu
  \phi.\lambda \ol{\alpha}.H)\ol{\beta}} \Eq_\rho[\ol{\beta:=
    R}]{\color{red}[\ol{\gamma :=S}]}
\end{array}\]

>>>>

Definition~\ref{def:wftypes} allows the formation of all of the types
from Section~\ref{sec:intro}:
\[\begin{array}{lll}
\mathit{List}\, \alpha & = & \mu \beta. \,\onet + \alpha \times
\beta\; = \; (\mu \phi. \lambda \beta.\,\onet + \beta \times \phi
\beta)\,\alpha\\ 
\mathit{PTree}\,\alpha & = & (\mu \phi. \lambda \beta.\,\beta +
\phi\,(\beta \times \beta))\,\alpha\\
\mathit{Forest}\,\alpha & = & (\mu \phi. \lambda \beta. \,\onet +
\beta \times \mathit{PTree}\,(\phi \beta))\,\alpha\\ 
\mathit{Bush}\,\alpha & = & (\mu \phi.\lambda \beta. \,\onet + \beta
\times \phi\,(\phi\beta))\,\alpha 
\end{array}\]

Note that since the body F of a type $(\mu \phi. \lambda
\ol\alpha. F)\ol{G}$ can only be functorial in $\phi$ and the
variables in $\ol{\alpha}$, the representation of
$\mathit{List}\,\alpha$ as the ADT $\mu \beta. \,\onet + \alpha \times
\beta$ cannot be functorial in $\alpha$. By contrast, if
$\mathit{List}\,\alpha$ is represented as the nested type $(\mu
\phi. \lambda \beta.\,\onet + \beta \times \phi \beta)\,\alpha$ then
we can choose $\alpha$ to be a functorial variable or not when forming
the type. This observation holds for other ADTs as well; for example,
if $\mathit{Tree}\,\alpha\,\gamma = \mu \beta. \alpha + \beta \times
\gamma \times \beta$, then $\alpha, \gamma; \emptyset \vdash
\mathit{Tree}\,\alpha\,\gamma$ is well-formed, but $\emptyset; \alpha,
\gamma \vdash \mathit{Tree}\,\alpha\,\gamma$ is not. And it also
applies to some non-ADT types, such as $\mathit{GRose}\,\phi\,\alpha =
\mu \beta. \onet + \alpha \times \phi\beta$, in which $\phi$ and
$\alpha$ must both be non-functorial variables. {\color{red} It is in
  fact possible to allow ``extra'' $0$-ary functorial variables in the
  body of $\mu$-types (functorial variables of higher arity are the
  real problem), which would allow the first-order representations of
  ADTs to be functorial.  However, it simplifies the formation rule
  for $\mu$-types considerably to disallow even these $0$-ary
  functorial variables. Since we can always use an ADT's (semantically
  equivalent) second-order representation when functoriality is
  needed, this restriction allows us to simplify the type syntax
  without sacrificing the expressivity of the calculus.}
%The representation of $\mathit{List}\,\alpha$ as the ADT $\mu
%\beta. \onet + \alpha \times \beta$ cannot be functorial in
%$\alpha$. By contrast, its representation as the nested type $(\mu
%\phi. \lambda \beta.\,\onet + \beta \times \phi \beta)\,\alpha$ can be
%considered either functorial in $\alpha$ or not, according to whether
%$\alpha$ is in the functorial or non-functorial context in which this
%type is formed. If $\emptyset;\alpha \vdash \mathit{List} \,\alpha$,
%then $\vdash \Nat^\alpha \onet\, (\mathit{List}\,\alpha)$ is
%well-formed; if $\alpha;\emptyset \vdash \List \,\alpha$, then it is
%not.  Similarly, if $\mathit{Tree}\,\alpha\,\gamma = \mu \,\beta.\,
%\alpha + \beta \times \gamma \times \beta$, then $\gamma;\emptyset
%\vdash \Nat^\alpha (\mathit{List}\, \alpha) \, (\mathit{Tree} \,
%\alpha\, \gamma)$ is well-formed, but $\emptyset;\gamma \vdash
%\Nat^\alpha (\mathit{List}\, \alpha) \, (\mathit{Tree} \, \alpha\,
%\gamma)$ is not. The body $F$ of a type $(\mu \phi.\lambda
%\ol\alpha.F)\ol G$ can only be functorial in $\phi$ and the variables
%in $\ol\alpha$. As a result, both $\phi$ and the variables in
%$\ol\alpha$ must be non-functorial in $\mathit{GRose}\, \phi \, \alpha
%= \mu\beta . \alpha \times \phi \beta$, $\mathit{GRose}\,\phi\,\alpha$
%cannot be the (co)domain of a $\Nat$-type binding any variables, and
%no type of the form $(\mu \phi. \lambda
%\alpha. \mathit{GRose}\,\phi\,\alpha)\tau$ is well-formed in our
%system.
These restrictions ensure that the set and relational interpretations
of types in Section~\ref{sec:type-interp} are functors. In
Section~\ref{sec:conclusion} we discuss how more expressive types
might be accommodated.

@@

\lambda
d{\color{red}\,\ol{B}\,\ol{C}.\,(}\mathit{in}_{{T}^X_{H,\rho{\color{red}[\ol{\gamma
          := 
        C}]}}}{\color{red})_{\ol{B}}} \;\;\mbox{ where } X \mbox{ is } \set \mbox{ when } \\ 

@@

\lambda d\,\eta{\color{red}\,\ol{B}\,\ol{C}.\,
(}\mathit{fold}_{T^X_{H,\rho{\color{red}[\ol{\gamma := C}]}}} \, (\lambda
\ol{A}.\,\eta_{\ol{A}\,\ol{C}}){\color{red})_{\ol{B}}}\\ 

@@

\section{Naturality and the Abstraction Theorem}\label{sec:free-theorems}

@@

We first show that terms of $\Nat$-type behave as natural
transformations with respect to their source and target functorial
types. If $\Gamma; \ol{\alpha} \vdash F$ then define the {\em
  identity} $\id_F$ on $F$ by $\id_F = \Gamma;\emptyset~|~\emptyset
\vdash L_{\ol{\alpha}}x.x : \Nat^{\ol{\alpha}} F\,F$, and if $\Gamma;
\emptyset \,|\, \Delta \vdash t: \Nat^{\overline{\alpha}} F\,G$ and
$\Gamma; \emptyset \,|\, \Delta \vdash s: \Nat^{\overline{\alpha}}
G\,H$ are terms then define the {\em composition} $s \circ t$ of $t$
and $s$ by $s \circ t = \Gamma; \emptyset\,|\, \Delta \vdash
L_{\overline{\alpha}}
x. s_{\overline{\alpha}}(t_{\overline{\alpha}}x):
\Nat^{\overline{\alpha}} F\,H$. Then \[\setsem{\Gamma; \emptyset \,|\,
  \emptyset \vdash \id_{F} : \Nat^{\ol{\alpha}} F\,F} \rho\, \ast =
\id_{\lambda \ol{A}. \setsem{\Gamma; \ol{\alpha} \vdash F} \rho
  [\ol{\alpha := A}]}\] for any set environment $\rho$, and
\[\setsem{\Gamma; \emptyset \,|\, \Delta \vdash s \circ t:
  \Nat^{\overline{\alpha}} F\,H} = \setsem{\Gamma; \emptyset \,|\,
  \Delta \vdash s: \Nat^{\overline{\alpha}} G\,H} \circ
\setsem{\Gamma; \emptyset \,|\, \Delta \vdash t:
  \Nat^{\overline{\alpha}} F\,G}\]
Naturality of term interpretations is then easily verified:
\begin{thm}\label{eq:ft-from-nat} 
  If\, $\Gamma; \emptyset \,|\, \Delta \vdash s : \Nat^{\overline{\alpha},
  \overline{\gamma}} F\,G$ and $\overline{\Gamma; \emptyset \,|\,
  \Delta \vdash t : \Nat^{\overline{\gamma}} K\, H}$, then
  \[\begin{array}{l}
  \hspace*{0.135in}
\setsem{\Gamma; \emptyset\,|\, \Delta
  \vdash 
  ((\map_G^{\overline{K}, \overline{H}})_\emptyset \,\overline{t}) \circ
(L_{\overline{\gamma}} z. s_{\overline{K}, \overline{\gamma}}
  z)
  : \Nat^{\overline{\gamma}} F[\overline{\alpha := K}]\,
  G[\overline{\alpha := H}]}\\
= \setsem{ \Gamma; \emptyset \,|\, \Delta \vdash
(L_{\overline{\gamma}} z.
  s_{\overline{H}, \overline{\gamma}} z)\circ
  ((\map_F^{\overline{K}, \overline{H}})_\emptyset \,
  \overline{t})  : \Nat^{\overline{\gamma}} F[\overline{\alpha :=
      K}]\, G[\overline{\alpha := H}]}
\end{array}\]
\end{thm}
\noindent

@@

\item
$\underline{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
  \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
    {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
  \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}$\; To
  see that if $d : \setsem{\Gamma;\emptyset \vdash \emptyset} \rho$
  then $$\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,
  \rho\,d$$ 
  is in $\setsem{\Gamma;\emptyset \vdash
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,
  \rho$, we first note that, for all $\ol{B}$ {\color{red} and $\ol{C}$},
  \begin{align*}
    &(\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,
  \rho\,d)_{\ol{B}{\color{red}\,\ol{C}}}\, \\
    =\,
    &(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}})_{\ol{B}}
  \end{align*}
  maps
  \begin{align*}
    &\setsem{\Gamma;\ol{\beta}{\color{red},\ol{\gamma}} \vdash H[\phi := (\mu
      \phi.\lambda {\overline \alpha}.H){\overline \beta}][\ol{\alpha
        := \beta}]}\rho[\ol{\beta := B}]{\color{red}[\ol{\gamma := C}]} 
        \\ = \,
    &T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}\, (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma
        := C}]}}) \, \ol{B}
  \end{align*}
        to $\setsem{\Gamma;\ol{\beta},{\color{red}\ol{\gamma}}
    \vdash (\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}
  \rho[\ol{\beta := B}]{\color{red}[\ol{\gamma := C}]} = (\mu
  T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}) \, \ol{B}$. 
  Secondly, we observe
  that
  \begin{align*}
    &\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
    \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
      {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
    \;(\mu \phi.\lambda {\overline \alpha}.H){\overline
      \beta}}\,\rho\,d \\ 
      = \, &\lambda
  \ol{B}{\color{red}\,\ol{C}}.\,(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma :=
          C}]}}})_{\ol{B}}
  \end{align*}
  is natural in $\ol{B}$ {\color{red} and $\ol{C}$}, since
{\color{red}  naturality of $\mathit{in}$ with respect to its functor argument and}
  naturality of $\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}}$ ensures
  that the following diagram commutes for all $\ol{f : B \to B'}$
  {\color{red} and
  $\ol{g : C \to C'}$}:
  \[\hspace*{-0.15in}\begin{tikzcd}[column sep=2.5in, row sep=0.75in]
%T^\set_{H,\rho[\ol{\gamma := C}]}\, (\mu T^\set_{H,\rho[\ol{\gamma := C}]})\, \ol{B}
%\ar[d, "{T^\set_{H,\id_\rho[\ol{\gamma := g}]}\,(\mu
%    T^\set_{H,\id_\rho[\ol{\gamma := g}]}) \, \ol{B}}"] \ar[r,
%  "{(\mathit{in}_{T^\set_{H,\rho[\ol{\gamma := C}]}})_{\ol{B}}}" ]
%& (\mu T^\set_{H,\rho[\ol{\gamma := C}]})\, \ol{B} \ar[d, "{(\mu
%    T^\set_{H,\id_\rho[\ol{\gamma := g}]}) \, \ol{B}}"]\\
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}\, (\mu
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\, \ol{B} \ar[d, "{T^\set_{H,\rho{\color{red}[\ol{\gamma :=
          C'}]}}\, (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\, \ol{f}}"]
\ar[r,"{(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})_{\ol{B}}}}" ] & 
\ar[d,"{\mu {T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}}\, {\ol{f}}}" ] 
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}\, (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\,
\ol{B'} \ar[r, "{(\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma :=
            C'}]}}})_{\ol{B'}}}"] & (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}})\,
\ol{B'}
\end{tikzcd}\]
}
  That, for all $\rho : \setenv$ and $d :
\setsem{\Gamma;\emptyset \vdash \emptyset}\rho$,
\[\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash
  \tin_H : \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
    {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
  \;(\mu \phi.\lambda {\overline \alpha}.H){\overline
    \beta}}\,\rho\,d\] satisfies the additional property needed for
it to be in 
\[\setsem{\Gamma;\emptyset \vdash
  \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
    {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
  \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}\,\rho\]
follows from the fact that, for every $\ol{R : \rel(B,B')}$
{\color{red} and $\ol{S : \rel(C,C')}$},
\[\hspace*{-0.2in}
  \begin{array}{ll}
 & (\,(\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
      \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
        {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
      \;(\mu \phi.\lambda {\overline \alpha}.H){\overline
        \beta}}\,\rho\,d)_{\ol{B}{\color{red},\ol{C}}},\,\\ & \hspace*{0.5in}(\setsem{\Gamma;\emptyset
      \,|\, \emptyset \vdash \tin_H : \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \,
      H[\phi := (\mu \phi.\lambda {\overline \alpha}.H){\overline
          \beta}][\ol{\alpha := \beta}] \;(\mu \phi.\lambda {\overline
        \alpha}.H){\overline
        \beta}}\,\rho\,d)_{\ol{B'}{\color{red},\ol{C'}}}\,)\\ &= (\,
    (\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}})_{\ol{B}},
    (\mathit{in}_{T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}})_{\ol{B'}}\,)
& (\, T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}\, (\mu
T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}) \, 
\ol{B} \to (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C}]}}) \, \ol{B}, \, \\
& \hspace*{0.5in}T^\set_{H,\rho{\color{red} [\ol{\gamma := C'}]}}\, (\mu
T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}) \, 
\ol{B'} \to (\mu T^\set_{H,\rho{\color{red}[\ol{\gamma := C'}]}}) \,\ol{B'} \, )\\
= &
 \relsem{\Gamma;\ol{\beta},{\color{red}\ol{\gamma}} \vdash H[\phi := (\mu
    \phi.\lambda {\overline \alpha}.H){\overline \beta}][\ol{\alpha :=
      \beta}]}\Eq_\rho[\ol{\beta := R}]{\color{red}[\ol{\gamma := S}]} \to\\
 & \hspace*{0.5in} \relsem{\Gamma;\ol{\beta}{\color{red},\ol{\gamma}} \vdash (\mu
  \phi.\lambda \ol{\alpha}.H)\ol{\beta}} \Eq_\rho[\ol{\beta:=
    R}]{\color{red}[\ol{\gamma :=S}]}
Finally, since $\Phi = \emptyset$, naturality of
\setsem{\Gamma;\emptyset \,|\, \emptyset \vdash \tin_H :
  \Nat^{\ol{\beta}{\color{red},\ol{\gamma}}} \, H[\phi := (\mu \phi.\lambda
    {\overline \alpha}.H){\overline \beta}][\ol{\alpha := \beta}]
  \;(\mu \phi.\lambda {\overline \alpha}.H){\overline \beta}}
